---
title: "Tracking Weather Patterns with 'riem'"
author: "Kathleen Scopis"
date: "`r Sys.Date()`"
TOC: true
code folding: hide
output: html_document
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Planners and public decision makers often rely on analyses relating to
geography and the natural environment. From predicting train delays to
addressing climate change, it is important to know historic weather
patterns for an area in order to better make predictions for the future.
The *'riem'* package helps to select ASOS stations (more on that below)
and report on the area's historical weather-related sensor readings. The
package is clean and simple to use, able to pull holistic data with just
a few lines of code.

## Step 1: Loading 'riem' Package

Before you can use the package's functions to pull data, it must first
be loaded into the environment. Using *install.packages()* and
*library()*, install the *riem* package and activate it in the project's
library.

```{r 1}
#install.packages('riem')
library(riem)

# other libraries needed for plotting
library(tidyverse)
library(ggplot2)
```

## Step 2: ASOS Network List

*'riem'* pulls information from the **Automated Surface/Weather
Observing Systems (ASOS)**. In order to select from the correct station,
first access the list of ASOS networks using the *riem_networks()*
command.

```{r 2}
riem_networks()
```

## Step 3: ASOS Station List

For this example, we will be observing sensor readings from Richmond, a
city in Virginia. To acquire the station ID, input the corresponding
network code obtained from running *riem_networks()*. According to the
last step, Virginia's corresponding ASOS network code is 'VA_ASOS'.

```{r 3}
riem_stations(network = 'VA_ASOS')
```

## Step 4: Pull Station Data

Once the station ID is selected, use *riem_measures()* to obtain
historical weather data. The syntax is as follows:

--- ***station:*** station ID

--- ***date_start:*** starting date of the desired data, typically in
YYYY-MM-DD format

--- ***date_end:*** ending date of the desired data, typically in
YYYY-MM-DD format

The resulting dataset features several columns, each representing a
different quality observed by the sensors. The data may include:

--- ***station***: three or four character site identifier

--- ***valid:*** timestamp of the observation (in UTC)

--- ***tmpf:*** Air Temperature (F), typically \@ 2 meters

--- ***dwpf***: Dew Point Temperature (F), typically \@ 2 meters

--- ***relh***: Relative Humidity in

--- ***drct***: Wind Direction in degrees from north

--- ***sknt***: Wind Speed in knots riem_measures 3

--- ***p01i***: One hour precipitation for the period from the
observation time to the time of the previous hourly precipitation reset.
This varies slightly by site. Values are in inches. This value may or
may not contain frozen precipitation melted by some device on the sensor
or estimated by some other means. Unfortunately, we do not know of an
authoritative database denoting which station has which sensor. • alti:
Pressure altimeter in inches

--- ***mslp***: Sea Level Pressure in millibar

--- ***vsby***: Visibility in miles

--- ***gust***: Wind Gust in knots

--- ***skyc1***: Sky Level 1 Coverage

--- ***skyc2***: Sky Level 2 Coverage

--- ***skyc3:*** Sky Level 3 Coverage

--- ***skyc4***: Sky Level 4 Coverage

--- ***skyl1:*** Sky Level 1 Altitude in feet

--- ***skyl2:*** Sky Level 2 Altitude in feet

--- ***skyl3***: Sky Level 3 Altitude in feet

--- ***skyl4:*** Sky Level 4 Altitude in feet

--- ***presentwx:*** Present Weather Codes (space seperated), see e.g.
Chapter 8 of [this
manual](https://www.ofcm.gov/publications/fmh/FMH1/FMH1.pdf) for further
explanations.

--- ***feel***: Apparent Temperature (Wind Chill or Heat Index) in degF

--- ***ice_accretion_1hr***: Ice Accretion over 1 Hour in inch

--- ***ice_accretion_3hr***: Ice Accretion over 3 Hour in inch

--- ***ice_accretion_6hr:*** Ice Accretion over 6 Hour in inch

--- ***elh:*** Relative Humidity in

--- ***metar***: unprocessed reported observation in METAR format

--- ***peak_wind_gust***: Wind gust in knots from the METAR PK WND
remark, this value may be different than the value found in the gust
field. The gust field is derived from the standard METAR wind report.

--- ***peak_wind_drct***: The wind direction in degrees North denoted in
the METAR PK WND remark.

--- ***peak_wind_time***: The timestamp of the PK WND value in the same
timezone as the valid field and controlled by the tz parameter.

# Example Analysis

Using the *riem* package, the following code pulls one week of
historical weather data for Richmond, VA, then graphs a few significant
variables.

```{r 4}
# pull data
weather.Data <- riem_measures(
  station = "OFP",
  date_start = "2023-09-17",
  date_end = "2023-09-23") %>% 
  filter(!is.na(tmpf)) # remove NAs

# plot data
par(mfrow = c(2,2))
(ggplot(weather.Data, aes(valid, tmpf))
    + geom_line()
    + labs(title = "Temperature",
           subtitle = "Richmond, VA",
           x = "date",
           y = "temperature (F)")
    )
(ggplot(weather.Data, aes(valid, relh))
    + geom_line()
    + labs(title = "Relative Humidity",
           subtitle = "Richmond, VA",
           x = "date",
           y = "humidity (%)")
    )
(ggplot(weather.Data, aes(valid, sknt))
    + geom_line()
    + labs(title = "Wind Speed",
           subtitle = "Richmond, VA",
           x = "date",
           y = "speed (knots)")
    )
(ggplot(weather.Data, aes(valid, vsby))
    + geom_line()
    + labs(title = "Visibility",
           subtitle = "Richmond, VA",
           x = "date",
           y = "visibility (miles)")
    )


```
